---
title: "plot_results"
output: pdf_document
date: "2024-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(ggrepel)
library(lubridate)
library(tidyr)
library(tidyverse)
```
# Read CSVS
```{r}
year=2023
reliability=read.csv("outage_records/metrics_2022.csv")
svi=read.csv("plots_and_metric_data/svi_interactive_map_2022.csv")
county_fips=read.csv('county_to_fips.csv')
wind_speed_customers=read.csv("processed_wind_customer_outage.csv")
county_customers=read.csv('mcc.csv')
EIA_disturbances<-read.csv("EIA_disturbances_data/all_illinois.csv")
EIA_saidi_saifi=read.csv('energy_information_administration_data/2022/Illinois/Reliability_2022_Illinois.csv')
service_territory=read.csv('energy_information_administration_data/2022/Illinois/Service_Territory_2022_Illinois.csv')
```

# Plot Reliability Metrics Against Social Vulnerability.
```{r}
reliability %>%
  ggplot(.,aes(x=saidi,y=svi))+geom_point()
reliability<-left_join(reliability,county_fips,by="county")
```

# Plot wind speed against percentage of customers out
```{r}
wind_speed_customers <- wind_speed_customers %>%
  mutate(date=as.Date(date,"%m/%d/%Y")) %>%
  mutate(year=as.integer(format(as.Date(wind_speed_customers$date,format="%m/%d/%Y"),"%Y")))
county_customers$County_FIPS=as.numeric(county_customers$County_FIPS)
wind_speed_customers <- left_join(wind_speed_customers,county_fips,by="county")
county_fips <- left_join(county_fips,county_customers,join_by(fips_code==County_FIPS))
wind_speed_customers <- left_join(wind_speed_customers,county_customers,join_by(fips_code==County_FIPS))
```

# Print the counties with the most at-risk reliability metrics.
```{r}
reliability %>%
  group_by(county) %>%
  summarise(saidi=(saidi),svi=svi) %>%
  top_n(10,saidi)
```
# SAIDI vs SAIFI
```{r}
 reliability%>%
  group_by(county) %>%
  mutate(SAIDI=mean(saidi),SAIFI=mean(saifi),svi=mean(svi))%>%
  ggplot(.,aes(x=SAIDI/60,y=SAIFI,color=svi))+geom_point()+geom_label(label=reliability$county,size=2)
```
# Convert EIA disturbances metrics into county-based format for SAIDI, SAIFI, and CAIDI.
```{r}
EIA_saidi_saifi$SAIDI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..SAIDI..minutes.per.year..)
EIA_saidi_saifi$SAIFI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..SAIFI..times.per.year..)
EIA_saidi_saifi$CAIDI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..CAIDI..minutes.per.interruption..)
service_reliability_mapper<-inner_join(service_territory,EIA_saidi_saifi,join_by('X.Utility.Name.'=='X.Utility.Characteristics..Utility.Name.'))
eia_reliability_county <- service_reliability_mapper %>%
  filter(!is.na(SAIDI)) %>%
  group_by(X.County.) %>%
  summarise(eia_SAIDI=mean(SAIDI),
         eia_SAIFI=mean(SAIFI),
         eia_CAIDI=mean(CAIDI)) %>%
  arrange(X.County.)
```
# Compare SAIDI, SAIFI, and CAIDI values between EIA database and EAGLE-I Calculations.
```{r}

residual_cmp<-left_join(reliability,eia_reliability_county,join_by(county==X.County.))
residual_cmp<-left_join(residual_cmp,svi,join_by(fips_code==FIPS))
residual_cmp <- residual_cmp %>%
  mutate(saidi_res=(eia_SAIDI-saidi),saifi_res=(eia_SAIFI-saifi),caidi_res=(eia_CAIDI-caidi))
residual_cmp %>%
  ggplot(.,aes(x=saidi,y=eia_SAIDI))+geom_point()+xlab("EAGLE-I SAIDI")+ylab("EIA SAIDI")+geom_abline(slope=1,intercept=0)
residual_cmp %>%
  ggplot(.,aes(x=saifi,y=eia_SAIFI))+geom_point()+xlab("EAGLE-I SAIFI")+ylab("EIA SAIFI")+geom_abline(slope=1,intercept=0)
residual_cmp %>%
  ggplot(.,aes(x=caidi,y=eia_CAIDI))+geom_point()+xlab("EAGLE-I CAIDI")+ylab("EIA CAIDI")+geom_abline(slope=1,intercept=0)
residual_cmp %>%
  ggplot(.,aes(x=saidi_res,y=saifi_res,label=county))+geom_point()+geom_text_repel()
```
# Continued comparison, but by utility instead of county
```{r}
service_reliability_mapper_eaglei=inner_join(service_reliability_mapper,reliability,join_by(X.County.==county))
service_reliability_mapper_eaglei <- service_reliability_mapper_eaglei %>%
  group_by(X.Utility.Name.)%>%
  summarise(county=X.County.,utility=X.Utility.Name.,saifi=mean(saifi,na.rm=TRUE),saidi=mean(saidi,na.rm=TRUE),caidi=mean(caidi,na.rm=TRUE))
EIA_service_util<-left_join(eia_reliability_county,service_reliability_mapper_eaglei,join_by(X.County.==county))
EIA_service_util<- EIA_service_util %>%
  mutate(diff_saidi=eia_SAIDI-saidi,diff_saifi=eia_SAIFI-saifi,diff_caidi=eia_CAIDI-caidi)
EIA_service_util %>%
  group_by(utility)%>%
  summarise(diff_saidi=mean(diff_saidi),diff_saifi=mean(diff_saifi),diff_caidi=mean(diff_caidi),utility=utility)
```


# Create a new dataframe that aggregates all outages by county into a single outage event.
```{r}
grouped_outages <- aggregate(wind_speed_customers$max_num_people_out,by=list(wind_speed_customers$date),FUN=sum)
grouped_customers <-aggregate(wind_speed_customers$Customers,by=list(wind_speed_customers$date),FUN=sum)
grouped_customer_minutes <- aggregate(wind_speed_customers$customer_mins,by=list(wind_speed_customers$date),FUN=mean)
grouped_customer_duration <-aggregate(wind_speed_customers$duration,by=list(wind_speed_customers$date),FUN=mean)
wind_speed <- aggregate(wind_speed_customers$wind_speed_mph,by=list(wind_speed_customers$date),FUN=mean)
wind_customers_grouped<-data.frame(id=grouped_outages$Group.1, num_outages=grouped_outages$x,                   num_customers=grouped_customers$x,wind_speed=wind_speed$x,customer_mins=grouped_customer_minutes$x, duration=grouped_customer_duration$x)
wind_customers_grouped<-wind_customers_grouped %>%
   mutate(pct_customers_out=num_outages/num_customers)
```

# EXPONENTIAL REGRESSION MODEL: Average Percent Customers Out (per wind speed)
```{r}
avg_wind_customers_grouped <- wind_customers_grouped %>%
  mutate(wind_speed=as.integer(round(wind_speed*1.15078))) %>%
  group_by(wind_speed) %>%
  mutate(avg_pct_out=mean(pct_customers_out),avg_cust_min=mean(customer_mins/num_outages))
avg_wind_customers_grouped_lm=lm(log(avg_wind_customers_grouped$avg_pct_out)~wind_speed,avg_wind_customers_grouped)
summary(avg_wind_customers_grouped_lm)
par(mfrow=c(2,2))
plot(avg_wind_customers_grouped_lm)
par(mfrow=c(1,1))
avg_wind_customers_grouped %>%
  ggplot(.,aes(x=wind_speed,y=avg_pct_out))+geom_point()+labs(x='Wind Speed (MPH)',y='Average Percent of Customers Out')
```
# Wind Speed v.s. Customer-Minutes
```{r}
avg_wind_customers_grouped %>%
  ggplot(.,aes(x=wind_speed,y=avg_cust_min))+geom_point()+labs(x='Wind Speed (MPH)',y='Average Customer Minutes Out')
```
# Add EIA disturbances to dataset
```{r}
EIA_disturbances_fmtd <- EIA_disturbances %>%
  mutate(id_start=parse_date_time(Date.Event.Began,orders=c("%m/%d/%Y","%Y-%m-%d","%Y-%m-%d %H%M%S")),
         id_end=parse_date_time(Date.of.Restoration,orders=c("%m/%d/%Y","%Y-%m-%d","%Y-%m-%d %H%M%S")))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  mutate(duration=as.numeric(difftime(strptime(paste(id_end,Time.of.Restoration),format="%Y-%m-%d   %H:%M:%S"),strptime(paste(id_start,Time.Event.Began),format="%Y-%m-%d %H:%M:%S"))))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  separate(.,Area.Affected,sep=c(":"),into=c("State","County"))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  mutate(customer_mins=duration*60*as.numeric(Number.of.Customers.Affected))

```

