---
title: "plot_results"
output: pdf_document
date: "2024-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r libraries}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(ggrepel)
library(lubridate)
library(tidyr)
library(tidyverse)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
library(sf)
```
# Read CSVS
```{r} 
year=2023
# Data by year
reliability=read.csv(paste0("outage_records/metrics_",year,".csv"))
svi=read.csv(paste0("plots_and_metric_data/svi_interactive_map_",year,".csv"))
EIA_saidi_saifi=read.csv(paste0('energy_information_administration_data/',year,'/Illinois/Reliability_',year,'_Illinois.csv'))
service_territory=read.csv(paste0('energy_information_administration_data/',year,'/Illinois/Service_Territory_',year,'_Illinois.csv'))
# all data
county_fips=read.csv('county_to_fips.csv')
wind_speed_customers=read.csv("processed_wind_customer_outage.csv")
county_customers=read.csv('mcc.csv')
EIA_disturbances<-read.csv("EIA_disturbances_data/all_illinois.csv")
geojson<-read_sf('illinois-with-county-boundaries_1097.geojson')
```
# Plot average outage duration
```{r}
ggplot(geojson)+geom_sf(aes(fill=reliability$saidi/60))+theme_void()+scale_fill_viridis_c(direction=1,name='Average Outage Duration (hrs)')
ggplot(geojson)+geom_sf(aes(fill=reliability$saifi))+theme_void()+scale_fill_viridis_c(direction=1,name='Average Outage Frequency')
```

# Select SVI features
```{r}
svi <- svi%>%
  select(COUNTY,FIPS,AREA_SQMI,E_TOTPOP,RPL_THEME1,RPL_THEME2,RPL_THEME3,RPL_THEME4,RPL_THEMES)
```

# Plot wind speed against percentage of customers out
```{r}
wind_speed_customers <- wind_speed_customers %>%
  mutate(date=as.Date(date,"%m/%d/%Y")) %>%
  mutate(year=as.integer(format(as.Date(wind_speed_customers$date,format="%m/%d/%Y"),"%Y")))
county_customers$County_FIPS=as.numeric(county_customers$County_FIPS)
wind_speed_customers <- left_join(wind_speed_customers,county_fips,by="county")
county_fips <- left_join(county_fips,county_customers,join_by(fips_code==County_FIPS))
wind_speed_customers <- left_join(wind_speed_customers,county_customers,join_by(fips_code==County_FIPS))
reliability <- left_join(reliability,county_fips,join_by(county==county))
reliability <- left_join(reliability,svi,join_by(fips_code==FIPS))
```
# SAIDI vs SAIFI
```{r}

reliability <- reliability %>%
  group_by(county) %>%
  mutate(SAIDI=mean(saidi)/60,SAIFI=mean(saifi),svi=mean(svi),pop_density=mean(E_TOTPOP/AREA_SQMI))
ggplot(reliability,aes(x=SAIDI,y=SAIFI,color=svi,size=pop_density))+geom_point()+labs(title="Outages by County, Illinois 2023",x="Average Outage Duration (hrs)",y="Average Outage Frequency",color="Social Vulnerability",size="Population Density")+theme_light()+geom_smooth(method="loess",span=2,show.legend=FALSE,color='darkgray')+scale_color_distiller(palette="YlGnBu",direction=1)
reliability_lm <- loess(reliability$SAIFI ~ reliability$SAIDI, data=reliability,span=2)
print(reliability_lm)
```
# Principal Component Analysis
```{r}
pca_matrix <- reliability%>%
  mutate(socioeconomic_status=RPL_THEME1,household_characteristics=RPL_THEME2,racial_ethnic_minority=RPL_THEME3,housing_type_transit=RPL_THEME4)
drops<-c("year","county","COUNTY","X","fips_code","AREA_SQMI","E_TOTPOP","RPL_THEMES","Customers","RPL_THEME1","RPL_THEME2","RPL_THEME3","RPL_THEME4","SAIDI","SAIFI")
pca_matrix <- pca_matrix[,!names(pca_matrix) %in% drops]
pca_matrix<-scale(pca_matrix)
corr <- round(cor(pca_matrix), 1)
ggcorrplot(corr,method="circle")+ggtitle("Correlation Matrix for Outage-Socioeconomic Data")
data.pca <- princomp(pca_matrix)
fviz_eig(data.pca, addlabels = TRUE)
fviz_pca_var(data.pca, col.var = "cos2",
            gradient.cols = c("yellow", "green", "blue"),
            repel = TRUE)
```

# Convert EIA disturbances metrics into county-based format for SAIDI, SAIFI, and CAIDI.
```{r}
EIA_saidi_saifi$SAIDI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..SAIDI..minutes.per.year..)
EIA_saidi_saifi$SAIFI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..SAIFI..times.per.year..)
EIA_saidi_saifi$CAIDI<-as.numeric(EIA_saidi_saifi$X.IEEE.Standard.All.Events..With.Major.Event.Days..CAIDI..minutes.per.interruption..)
service_reliability_mapper<-inner_join(EIA_saidi_saifi,service_territory,join_by('X.Utility.Characteristics..Utility.Name.'=='X.Utility.Name.'))
eia_reliability_county <- service_reliability_mapper %>%
  filter(!is.na(SAIDI),!is.na(SAIFI)) %>%
  group_by(X.County.) %>%
  summarise(utility=X.Utility.Characteristics..Utility.Name.,
            eia_SAIDI=(SAIDI),
         eia_SAIFI=(SAIFI),
         eia_CAIDI=(CAIDI)) %>%
  arrange(X.County.)
```
# Compare SAIDI, SAIFI, and CAIDI values between EIA database and EAGLE-I Calculations.
```{r}
residual_cmp<-left_join(reliability,eia_reliability_county,join_by(county==X.County.))
residual_cmp <- residual_cmp %>%
  mutate(saidi_res=(eia_SAIDI-saidi)/saidi,saifi_res=(eia_SAIFI-saifi)/saifi,caidi_res=(eia_CAIDI-caidi)/caidi)
rel_by_county <- residual_cmp %>%
  group_by(county) %>%
  summarise(utility=list(utility),
            eaglei_saidi=mean(saidi,na.rm=TRUE),
            eaglei_saifi=mean(saifi,na.rm=TRUE),
            eaglei_caidi=mean(caidi,na.rm=TRUE),
            eia_SAIDI=mean(eia_SAIDI,na.rm=TRUE),
            eia_SAIFI=mean(eia_SAIFI,na.rm=TRUE),
            eia_CAIDI=mean(eia_CAIDI,na.rm=TRUE),
            saidi_res=mean(saidi_res,na.rm=TRUE),
            saifi_res=mean(saifi_res,na.rm=TRUE),
            caidi_res=mean(caidi_res,na.rm=TRUE),
            pop_density=mean(pop_density,na.rm=TRUE),
            svi=mean(svi,na.rm=TRUE)) 
rel_by_county %>%
  ggplot(.,aes(x=saidi_res,y=saifi_res,label=county,color=svi,size=pop_density))+geom_point()+labs(x="% Difference in Outage Duration",y="% Difference in Outage Frequency",color="Social Vulnerability",size="Population Density")+ggtitle("Difference in Reported Outage Values, County")+scale_x_sqrt()+scale_y_sqrt()+theme_light()+geom_text_repel(size=3,color='darkslategray')+scale_color_distiller(palette="YlGnBu",direction=1)
rel_by_utility <- residual_cmp %>%
  group_by(utility) %>%
  summarise(
            eaglei_saidi=mean(saidi,na.rm=TRUE),
            eaglei_saifi=mean(saifi,na.rm=TRUE),
            eaglei_caidi=mean(caidi,na.rm=TRUE),
            eia_SAIDI=mean(eia_SAIDI,na.rm=TRUE),
            eia_SAIFI=mean(eia_SAIFI,na.rm=TRUE),
            eia_CAIDI=mean(eia_CAIDI,na.rm=TRUE),
            saidi_res=mean(saidi_res,na.rm=TRUE),
            saifi_res=mean(saifi_res,na.rm=TRUE),
            caidi_res=mean(caidi_res,na.rm=TRUE),
            pop_density=mean(pop_density,na.rm=TRUE),
            svi=mean(svi,na.rm=TRUE))
rel_by_utility %>%
  ggplot(.,aes(x=saidi_res,y=saifi_res,label=utility,color=svi,size=pop_density))+geom_point()+geom_text_repel(size=3,color='darkslategray')+labs(x="% Difference in Outage Duration",y="% Difference in Outage Frequency",color="Social Vulnerability",size="Population Density")+ggtitle("Difference in Reported Outage Values, Utility")+theme_light()+scale_color_distiller(palette="YlGnBu",direction=1)
```

# Create a new dataframe that aggregates all outages by county into a single outage event.
```{r}
grouped_outages <- aggregate(wind_speed_customers$max_num_people_out,by=list(wind_speed_customers$date),FUN=sum)
grouped_customers <-aggregate(wind_speed_customers$Customers,by=list(wind_speed_customers$date),FUN=sum)
grouped_customer_minutes <- aggregate(wind_speed_customers$customer_mins,by=list(wind_speed_customers$date),FUN=mean)
grouped_customer_duration <-aggregate(wind_speed_customers$duration,by=list(wind_speed_customers$date),FUN=mean)
wind_speed <- aggregate(wind_speed_customers$wind_speed_mph,by=list(wind_speed_customers$date),FUN=mean)
wind_customers_grouped<-data.frame(id=grouped_outages$Group.1, num_outages=grouped_outages$x,                   num_customers=grouped_customers$x,wind_speed=wind_speed$x,customer_mins=grouped_customer_minutes$x, duration=grouped_customer_duration$x)
wind_customers_grouped<-wind_customers_grouped %>%
   mutate(pct_customers_out=num_outages/num_customers)
```

# EXPONENTIAL REGRESSION MODEL: Average Percent Customers Out (per wind speed)
```{r}
avg_wind_customers_grouped <- wind_customers_grouped %>%
  mutate(wind_speed=round(wind_speed*1.15078/2)*2) %>%
  group_by(wind_speed) %>%
  mutate(avg_pct_out=mean(pct_customers_out),avg_out=mean(num_outages),avg_cust_min=mean(customer_mins/num_outages))
avg_wind_lm <- avg_wind_customers_grouped %>%
  filter(wind_speed<70)
avg_wind_customers_grouped_lm=lm(log(avg_wind_lm$avg_out)~wind_speed,avg_wind_lm)
summary(avg_wind_customers_grouped_lm)
avg_wind_customers_grouped %>%
  ggplot(.,aes(x=wind_speed,y=avg_out))+geom_col(fill="#41B7C4")+labs(x='Wind Speed (MPH)',y='Average Number of Customers Out')+theme_light()+ggtitle("Wind Speed vs. Number of Customers Affected in Illinois")+scale_x_continuous(breaks=seq(40,80,5),limits=c(40,80))+scale_y_continuous(labels=scales::label_comma())
```
```{r}
mat<-wind_customers_grouped
drops<-c("id","num_customers","avg_cust_min")
mat <- mat[,!names(wind_customers_grouped) %in% drops]
mat<-scale(mat)
corr <- round(cor(mat), 1)
ggcorrplot(corr,method="circle")+ggtitle("Correlation Matrix for Weather-Outage Data")
```

# Add EIA disturbances to dataset
```{r}
EIA_disturbances_fmtd <- EIA_disturbances %>%
  mutate(id_start=parse_date_time(Date.Event.Began,orders=c("%m/%d/%Y","%Y-%m-%d","%Y-%m-%d %H%M%S")),
         id_end=parse_date_time(Date.of.Restoration,orders=c("%m/%d/%Y","%Y-%m-%d","%Y-%m-%d %H%M%S")))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  mutate(duration=as.numeric(difftime(strptime(paste(id_end,Time.of.Restoration),format="%Y-%m-%d   %H:%M:%S"),strptime(paste(id_start,Time.Event.Began),format="%Y-%m-%d %H:%M:%S"))))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  separate(.,Area.Affected,sep=c(":"),into=c("State","County"))

EIA_disturbances_fmtd <- EIA_disturbances_fmtd %>%
  mutate(customer_mins=duration*60*as.numeric(Number.of.Customers.Affected))

```

